parasitemia

### 2023-06-12 Scripts/R/markdown_files/response_to_reviewers.Rmd

---
title: "Cerebral oxygenation manuscript response to reviewers"
date: "last updated: `r Sys.Date()`"
output: pdf_document
---

<style type="text/css">
  body{
  font-family: Arial;
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10, 
                      fig.align = 'center')

# Load libraries
library(tidyverse)
library(readxl)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)
library(psych)
library(patchwork)

# Load data
base_dir <- "/Users/ray/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Ackerman Lab 2019-2021/cerebral_oxygenation/"

# ALT LOADING: use base_dir
#load(paste0(base_dir, "objects/20221125_final_glm_models.Rdata"))
#load(paste0(base_dir, "objects/bandpass_filtered_dfa_results.Rdata"))
#load(paste0(base_dir, "objects/sims.Rdata"))

font_import()

# set plot theme
theme_set(theme_bw() +
              theme(plot.title = element_text(size = 15),
                    axis.title = element_text(size = 15),
                    axis.text = element_text(size = 15),
                    strip.text = element_text(size = 15),
                    legend.title = element_text(size = 15),
                    legend.text = element_text(size = 12)))

```

```{r data}

# read in data

df_hbox <- read_csv(paste0(base_dir, "Data/Hans_datatable_exports/malawi key data v04Jan2022.csv")) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

  dplyr::filter(status %in% c("CM", "HC", "UM") & session_no == 1 | subject_id == "TM0003CM01") %>% 
  dplyr::filter(!grepl("blood", subject_id)) %>% 
  mutate(status = factor(status, levels = c("HC", "UM", "CM"))) %>% 
  mutate_at(c("bp_dias", "bp_sys", "glucose", "lactate"), as.numeric) %>% # converts remaining character variables to numerics
  select(subject_id, status, everything()) %>% 
  group_by(status)

```

## Reviewer 2

### Minor comment 1

```{r r2c1, fig.height = 10, fig.width = 15}

# MEDIAN (25th, 75th)
df_hbox %>% 
  filter(status == "CM") %>% 
  ungroup %>% 
  select(hrp2, parasitemia) %>% 
  pivot_longer(1:ncol(.), names_to = "variable", values_to = "value") %>% 
  group_by(variable) %>% 
  summarize(median = median(value, na.rm = TRUE),
            `25th` = quantile(value, probs = 0.25, na.rm = TRUE), 
            `75th` = quantile(value, probs = 0.75, na.rm = TRUE)
  ) %>%

  knitr::kable()

# TEST FOR NORMALITY
features <- c("hrp2", "parasitemia", "cerebral_hb_tot", "muscle_hb_tot", "cerebral_hb_tot_alpha2", "muscle_hb_tot_alpha2")

df_normality <- df_hbox %>% 
  filter(status == "CM") %>% 
  ungroup %>% 
  select_at(features) %>% 
  pivot_longer(1:ncol(.), names_to = "feature", values_to = "value") %>%
  group_by(feature) %>% 
  mutate(sw_p = shapiro.test(value)$p.value,
         sw_p = ifelse(sw_p < 0.05, scientific(sw_p, digits = 3), as.character(round(sw_p, 2))),
         sw_p = paste0("SW p = ", sw_p),

         xmin = density(value, na.rm = TRUE)$x %>% min %>% floor,
         xmax = density(value, na.rm = TRUE)$x %>% max %>% ceiling,
         ymin = density(value, na.rm = TRUE)$y %>% min %>% floor,
         ymax = density(value, na.rm = TRUE)$y %>% max %>% ceiling

  )

df_text <- df_normality %>% 
  dplyr::select(-value) %>% 
  distinct() %>% 
  mutate(x = xmax - 0.1*xmax,
         y = ymax - 0.1*ymax) %>% 
  dplyr::select(-contains("min"), -contains("max")) %>% 
  dplyr::rename("label" = "sw_p")

p_sw <- df_normality %>% 
  ggplot(aes(x = value)) +
  geom_density(trim = FALSE) +
  geom_text(
    data = df_text,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1,
    vjust = 1,
    size = 5
  ) + 
  labs(y = "") +
  facet_wrap(vars(feature), scales = "free")

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND CEREBRAL HBTOT
p_hrp2_cerebral <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = log10(cerebral_hb_tot))) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "log10(cerebral Hbtot)") +
  ggtitle("NIRS-derived cerebral Hbtot and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_cerebral <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = log10(cerebral_hb_tot))) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "log10(cerebral Hbtot)") +
  ggtitle("NIRS-derived cerebral Hbtot and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND MUSCle HBTOT
p_hrp2_muscle <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = muscle_hb_tot)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  xlim(c(2.5, 4.5)) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = bquote(`Muscle Hb`[tot])) +
  ggtitle("NIRS-derived muscle Hbtot and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_muscle <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = muscle_hb_tot)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = bquote(`muscle Hb`[tot])) +
  ggtitle("NIRS-derived muscle Hbtot and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND CEREBRAL HBTOT ALPHA
p_hrp2_cerebral_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = cerebral_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "Cerebral Hbtot alpha") +
  ggtitle("NIRS-derived cerebral Hbtot alpha and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_cerebral_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = cerebral_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "Cerebral Hbtot alpha") +
  ggtitle("NIRS-derived cerebral Hbtot alpha and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# PLOT CORRELATIONS BETWEEN QUANTITATIVE PARASITEMIA MEASUREMENTS IN CM AND MUSCLE HBTOT ALPHA
p_hrp2_muscle_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(hrp2), y = muscle_hb_tot_alpha)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  xlim(c(2.5, 4.5)) +
  labs(x = paste0("log10(HRP2) (ng/\u03bcL)"), y = "Muscle Hbtot alpha") +
  ggtitle("NIRS-derived muscle Hbtot alpha and HRP2 in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

p_parasitemia_muscle_alpha <- df_hbox %>% 
  filter(status == "CM") %>%

  ggplot(aes(x = log10(parasitemia), y = muscle_hb_tot_alpha2)) +
  geom_point(size = 4, shape = 1) +
  geom_smooth(method = "lm") +
  stat_cor(size = 5) +
  labs(x = "log10(parasitemia) (RBCs/\u03bcL)", y = "Muscle Hbtot alpha") +
  ggtitle("NIRS-derived muscle Hbtot alpha and parasitemia in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

# COMBINE PLOTS

p_sw

(p_hrp2_cerebral | p_parasitemia_cerebral) /
(p_hrp2_muscle | p_parasitemia_muscle)

(p_hrp2_cerebral_alpha | p_parasitemia_cerebral_alpha) /
(p_hrp2_muscle_alpha | p_parasitemia_muscle_alpha)

```

## Reviewer 3

### Recommendation 1

```{r r3r1}

# REGRESSION
# df_hbox %>% 
#   filter(status == "CM") %>% 
#   
#   ggplot(aes(x = cerebral_hb_tot, y = outcome)) +
#   geom_point(size = 4, shape = 1) +
#   geom_smooth(method = "lm") +
#   stat_cor(size = 5) +
#   labs(x = "Cerebral Hbtot", y = "Outcome") +
#   ggtitle("Correlation between NIRS-derived cerebral Hbtot and outcome in children with CM") +
#   theme_bw() +
#   theme(axis.text = element_text(size = 15),
#         axis.title = element_text(size = 15))

# KRUSKAL-WALLIS
kw_p <- kruskal.test(cerebral_hb_tot ~ outcome,
             data = df_hbox %>%
               filter(status == "CM") %>% 
               mutate(outcome = as.factor(outcome))) %>% 
  .$p.value %>% 
  round(2)

df_hbox %>%
  filter(status == "CM") %>%
  mutate(outcome =

           case_when(

             outcome == 1 ~ "survival",
             outcome == 2 ~ "survival with neurological sequelae",
             outcome == 3 ~ "non-survival"

           ) %>%

           factor(levels = c("survival", "survival with neurological sequelae", "non-survival"))

           ) %>%

  ggplot(aes(x = outcome, y = cerebral_hb_tot, color = outcome)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 4, alpha = 0.5,
             position = position_jitter(width = 0.1)) +
  #annotate(geom = "text", label = paste0("KW p-value = ", kw_p), x = 3, y = 350, size = 5) +
  labs(x = "", y = bquote(`Cerebral Hb`[tot])) +
  ggtitle("Outcome and NIRS-derived cerebral Hbtot in children with CM") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.position = "none"
        )

```

### 2023-06-12 Scripts/R/markdown_files/20230109_fig1_fig3_reformat.Rmd
12 June 2023 clean and organize scripts for publishing

---
title: "Cerebral oxygenation manuscript figures"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

# Load libraries
library(tidyverse)
library(readxl)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)

# Load data
load("objects/glm_model.Rdata")
load("objects/bandpass_filtered_dfa_results.Rdata")
load("objects/sims.Rdata")

@@

### Table 1. Baseline clinical characteristics

#### Table 1a. All clinical data
```{r table1a, eval = FALSE}

# number of subjects 
df_status <- df_hbox %>% 
  count() %>% 
  pivot_wider(1:2, names_from = "status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  count(sex) %>% 
  dplyr::filter(sex != "") %>% 
  pivot_wider(names_from = "sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>%

  pivot_wider(c(1,4), names_from = "status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# other numeric variables

  # select variables for table
  clinical_voi <- c("age_calc", "temperature", "hr", "bp_sys", "bp_dias", "resp_rate", "pulse_ox", "hct", "lactate", "glucose", "aa_arg")

  # calculate median, first and third quartiles
  df_clinical_res <- df_hbox %>% 
    select_at(clinical_voi) %>% 
    summarise_all(funs(median = median, q1 = quantile(., probs = 0.25), q3 = quantile(., probs = 0.75)), na.rm = TRUE) %>% 
    mutate_if(is.numeric, round, 2)

  # write function to paste together values
  f_unite_cols <- function(col) {
    df_clinical_res %>% 
      unite(!! col, starts_with(col), sep = "_") %>% 
      dplyr::select(!! col) 
    }

  # paste together and transpose
  df_clinical_res <- 1:length(clinical_voi) %>% 
    purrr::map_dfc(~f_unite_cols(clinical_voi[.x])) %>% 
    mutate_all(funs(str_replace(., "_", " ("))) %>% 
    mutate_all(funs(str_replace(., "_", ", "))) %>% 
    mutate_all(funs(paste0(., ")"))) %>%

    t() %>% 
    as_tibble(rownames = "variable") %>% 
    dplyr::rename("HC" = "V1", "UM" = "V2", "CM" = "V3") %>% 
    add_row(df_status, .before = 1) %>% 
    add_row(df_sex, .before = 2)

# export to CSV
  write.csv(df_clinical_res, file = "Outputs/table_1.csv")

# create kable df
  df_clinical_res %>% knitr::kable(align = rep('c', 4))

```
**Table 1a.** A summary of subject information and hematological data. Group patient data presented as median (quartile 1, quartile 3).

#### Table 1b. CM-specific data
```{r table1b, eval = FALSE}

cm_voi <- c("bcs_session", "brain_swell", "hrp2")

df_cm_voi <- df_hbox %>% 
  ungroup %>% 
  filter(status == "CM") %>% 
  select_at(cm_voi) %>% 
  mutate_at(c("bcs_session", "brain_swell"), as.factor)

# BCS counts
df_cm_voi %>% count(bcs_session) %>% knitr::kable()

# Brain swell score counts
df_cm_voi %>% count(brain_swell) %>% knitr::kable()

# HRP
df_cm_hrp2 <- df_cm_voi %>% 
  select(cm_voi[3]) %>% 
  summarise_all(funs(median = median, q1 = quantile(., probs = 0.25), q3 = quantile(., probs = 0.75)), na.rm = TRUE) %>% 
  unite(hrp2, 1:2, sep = " (") %>% 
  unite(hrp2, 1:2, sep = ", ") %>% 
  mutate(hrp2 = paste0(hrp2, ")"))

df_cm_hrp2 %>% knitr::kable()

```
**Table 1b.** CM-specific clinical characteristics

@@

# extract hb_tot and hb_oxy for the brain and muscle
df_fig1 <- df_hbox %>% 
  select(contains(c("hb_tot","hb_oxy")), -contains(c("alpha", "bsl"))) %>%
  ungroup() %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("tissue", "variable"), sep = "_", extra = "merge")

@@

```
 \newpage

### Fig 2. Hb_oxy alpha and Hb_tot alpha in the brain and muscle
```{r fig2}

# extract hb_tot and hb_oxy for the brain and muscle
df_fig2 <- df_hbox %>% 
  select(contains("alpha2")) %>%
  ungroup() %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("tissue", "variable"), sep = "_", extra = "merge") %>% 
  mutate(variable = str_remove(variable, "2"))

# plot
df_fig2 %>%

  geom_hline(yintercept = 0.4, lty = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.6, lty = 2, alpha = 0.7) +
  geom_hline(yintercept = 1.0, lty = 2, alpha = 0.7) +
  geom_hline(yintercept = 1.5, lty = 2, alpha = 0.7) +

@@

```
 \newpage

### Fig 3. Increased cerebral_hb_tot and decreased cerebral_hb_tot_alpha are distinct clinical features of cerebral malaria
```{r fig3_data}

# select model variables of interest
model_voi <- c("subject_id", "status", 
               "hct", "lactate",
               "cerebral_hb_tot", "cerebral_hb_oxy",
               "cerebral_hb_tot_alpha2", "cerebral_hb_oxy_alpha2")

df_model <- df_hbox %>% 
  select_at(model_voi) %>%

  # remove HC
  ungroup %>% 
  filter(status != "HC")

# impute missing data with median by group
  f_impute <- function(x, na.rm = TRUE) (replace(x, is.na(x), median(x, na.rm = na.rm)))

  df_model_impute <- df_model %>% 
    group_by(status) %>% 
    mutate_at(3:ncol(.), f_impute) %>% 
    ungroup()

@@

#### A. Correlations of variables within subject groups
```{r fig3a, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

# calculate spearman correlations in each group

f_calc_cor <- function(df, group) {

  m <- df %>%
    dplyr::filter(status == group) %>% 
    dplyr::select(-c(subject_id, status)) %>% 
    cor(method = "spearman")

  diag(m) <- 0

  m
}

# create run correlations of potential regressors
m_cor_um <- df_model_impute %>% f_calc_cor(group = "UM")

m_cor_cm <- df_model_impute %>% f_calc_cor(group = "CM")

# correlation matrix

# write plot function
  f_plot_cor <- function(mat) {

  melt(mat) %>%

    mutate_if(is.factor, factor, levels = c("hct", "lactate", "cerebral_hb_oxy", "cerebral_hb_tot", "cerebral_hb_oxy_alpha2", "cerebral_hb_tot_alpha2")) %>%

    ggplot(aes(x = X1, y = X2, fill = value, label = ifelse(value != 0, round(value, 2), NA))) + 
    geom_tile() +
    scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
    geom_text(size = 9) +

    xlab("") +
    ylab("") +
    theme_bw() +
    theme(# text = element_text(family = "Arial"),
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text = element_text(size = 35),
          legend.text = element_text(size = 40),
          legend.title = element_text(size = 50),
          legend.key.size = unit(2, 'cm'),
          plot.title = element_text(size = 30)
    )

}

# plot

  # extract legend
  p_leg <- m_cor_um %>% f_plot_cor() %>% get_legend()

  # create plots
  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria") + theme(legend.position = "none")

  # as_ggplot(p_leg)

```
\newpage

#### B. Principal component analysis
```{r fig3b}

# convert tibble to df
df_model_pca <- df_model_impute %>% 
  dplyr::select(-c(subject_id, status)) %>% 
  as.data.frame()

# add rownames
rownames(df_model_pca) <- df_model_impute$subject_id

# compute PCA
my_pca <- prcomp(df_model_pca, scale = TRUE)

# define groups
groups <- as.factor(df_model_impute$status)

# Eigenvalues
pca_eig_val <- get_eigenvalue(my_pca)

# two data frames for plotting
df_vars <- my_pca$rotation %>% as_tibble() %>% mutate(variable = rownames(my_pca$rotation))
df_pnts <- my_pca$x %>% 
  as_tibble() %>% 
  mutate(status = groups, 
         subject_id = df_model_impute$subject_id)

# plot

  # plot patients
ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status), size = 3) +
  stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, fill = status),
                    level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  # stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, shape = status),
  #           na.rm = FALSE) +

  # View patient IDs
  # geom_text_repel(data = df_pnts, aes(x = PC1, y = PC2, label = subject_id, size = 10), show.legend = FALSE) +

  # plot variables
  # geom_point(data = df_vars, aes(x = PC1*5, y = PC2*5)) +
  geom_text_repel(data = df_vars, aes(x = PC1*5, y = PC2*5, label = variable, size = 10), show.legend = FALSE) +
  geom_segment(data = df_vars, aes(x = 0, y = 0, xend = PC1*5, yend = PC2*5), arrow = arrow()) +

  # draw lines
  geom_vline(xintercept = 0, lty = 2, color = "black") +
  geom_hline(yintercept = 0, lty = 2, color = "black") +

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC2: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +

  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

```
\newpage

#### C. Coefficients
```{r model, eval = FALSE}

### shows how logistic regression model was set up and run, results saved in objects/glm_model ###

# resample
set.seed(234)
hbox_boot <- bootstraps(df_model, strata = Status, times = 5000)

# recipe
hbox_rec <- recipe(Status ~ ., data = df_model) %>% 
  step_impute_knn(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_smote(Status) %>% 
  update_role(subject_id, new_role = "id")

# create workflow
hbox_wf <- workflow() %>% add_recipe(hbox_rec)

# logistic regression model specifications
glm_spec <- logistic_reg() %>% set_engine("glm")

# create model on 5000 different resamples
glm_res <- hbox_wf %>% 
  add_model(glm_spec) %>%
  fit_resamples(
    resamples = hbox_boot,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = tune::control_resamples(save_pred = TRUE, verbose = TRUE)
  )

# select best model to be used for evaluation
glm_best <- glm_res %>% select_best("roc_auc")

# finalize model
hbox_final <- hbox_wf %>% 
  add_model(glm_spec) %>% 
  finalize_workflow(glm_best) %>%
  fit(df_model) %>%
  extract_fit_parsnip()

```{r table3c}

df_CI <- confint(hbox_final$fit) %>% exp() %>% as_tibble(rownames = "term")

table3c <- hbox_final %>%
  tidy(exponentiate = TRUE) %>%
  filter(p.value < 0.05) %>% 
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  filter(Term != "(Intercept)") %>% 
  as.data.frame()

table3c %>% knitr::kable(align = rep('c', 5))

```
**note: for the OR, denominator is probability of patient having CM; numerator is p patient having UM

\newpage

#### D. ROC curve of final model
```{r fig3d_plot, fig.height = 8}

df_roc <- glm_res %>% 
  collect_predictions() %>% 
  roc_curve(status, .pred_CM)

ggplot() +
  geom_path(data = df_roc, 
            aes(x = 1 - specificity, y = sensitivity), 
            size = 2.5) +
  geom_abline(lty = 3) +
  labs(x = "False positive rate", y = "True positive rate") +
  theme_bw() +

  theme(# text = element_text(family = "Arial"),
        axis.title = element_text(size = 25),
        axis.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.position = "none")

```{r fig3d_inset}

fig3d_inset <- glm_res %>% 
  collect_metrics() %>% 
  dplyr::select(.metric, mean, std_err) %>% 
  mutate(mean = mean*100, std_err = std_err*100) %>% 
  mutate(Metric = c("Accuracy", "ROC AUC", "Sensitivity", "Specificity")) %>% 
  dplyr::select(c("Metric", "mean", "std_err")) %>%

  dplyr::rename("Mean" = "mean", "Standard error" = "std_err") %>% 
  mutate_if(is.numeric, round, 2) %>% 
  as.data.frame()

fig3d_inset %>% knitr::kable(align = rep('c', 4))

```
\newpage

### Fig 4. An increase in cerebral_hb_tot is significantly associated with an increase in brain swelling score
```{r brainswell_model_comparison, eval = FALSE}

# remove cerebral_hb_tot outliers
  df_hbox_hbtot_outliers <- df_hbox %>% 
    mutate(Q1_hbtot = quantile(cerebral_hb_tot, 0.25, na.rm = TRUE),
           Q3_hbtot = quantile(cerebral_hb_tot, 0.75, na.rm = TRUE),
           IQR_hbtot = Q3_hbtot - Q1_hbtot,
           outlier_hbtot = case_when(

             cerebral_hb_tot < Q1_hbtot - IQR_hbtot ~ 1,
             cerebral_hb_tot > Q3_hbtot + IQR_hbtot ~ 1,
             TRUE ~ 0

           )

    ) %>% 
    filter(outlier_hbtot != 1)

lm1 <- lm(brain_swell ~ cerebral_hb_tot, data = df_hbox_hbtot_outliers)
lm2 <- lm(brain_swell ~ cerebral_hb_tot + hct, data = df_hbox_hbtot_outliers)
lm3 <- lm(brain_swell ~ cerebral_hb_tot + cerebral_hb_tot_alpha2, data = df_hbox_hbtot_outliers)
lm4 <- lm(brain_swell ~ cerebral_hb_tot + cerebral_hb_tot_alpha2 + hct, data = df_hbox_hbtot_outliers)
lm5 <- lm(brain_swell ~ cerebral_hb_oxy + cerebral_hb_oxy_alpha2 + cerebral_hb_tot, data = df_hbox_hbtot_outliers)

# evaluate parameters for all models
brain_swell_models <- list(lm1, lm2, lm3, lm4, lm5)

df_brain_swell_models <- tibble(formula = as.character(),
                                term = as.character(),
                                aic = as.numeric(),
                                r2 = as.numeric(),
                                adj_r2 = as.numeric(),
                                model_p_val = as.numeric(),
                                estimate = as.numeric(),
                                std_error = as.numeric(),
                                statistic = as.numeric(),
                                term_p_val = as.numeric())

for (i in 1:length(brain_swell_models)) {

  formula <- brain_swell_models[[i]] %>% .$call
  formula <- gsub(".*=", "", formula)[[2]]

  aic <- brain_swell_models[[i]] %>% AIC
  r2 <- brain_swell_models[[i]] %>% summary %>% .$r.squared
  adj_r2 <- brain_swell_models[[i]] %>% summary %>% .$adj.r.squared
  f <- brain_swell_models[[i]] %>% summary %>% .$fstatistic
  p <- pf(f[1], f[2], f[3], lower.tail = FALSE)
  coef <- brain_swell_models[[i]] %>% tidy %>% filter(term != "(Intercept)") %>% dplyr::rename("term_p_val" = "p.value")

  df_tmp <- tibble(formula = formula,
                   term = coef$term,
                   aic = aic,
                   r2 = r2,
                   adj_r2 = adj_r2,
                   model_p_val = p) %>%
    left_join(coef) %>%
    clean_names

  df_brain_swell_models <- df_brain_swell_models %>% 
    bind_rows(df_tmp) %>% 
    relocate(term, .before = estimate)

}

df_brain_swell_models %>% write.csv(file = "outputs/brain_swell_models.csv")

```{r fig4_plot}

# select model voi
brain_swell_voi <- c("brain_swell", "cerebral_hb_tot", "hct")
df_fig4 <- df_hbox %>% select_at(brain_swell_voi)

# remove cerebral_hb_tot outliers
  df_fig4_outliers <- df_fig4 %>% 
    mutate(Q1_hbtot = quantile(cerebral_hb_tot, 0.25, na.rm = TRUE),
           Q3_hbtot = quantile(cerebral_hb_tot, 0.75, na.rm = TRUE),
           IQR_hbtot = Q3_hbtot - Q1_hbtot,
           outlier_hbtot = case_when(

             cerebral_hb_tot < Q1_hbtot - IQR_hbtot ~ 1,
             cerebral_hb_tot > Q3_hbtot + IQR_hbtot ~ 1,
             TRUE ~ 0

           )

    ) %>% 
    filter(outlier_hbtot != 1)

# model
  lm_fig4 <- lm(brain_swell ~ cerebral_hb_tot + hct, data = df_fig4_outliers)
  f <- summary(lm_fig4)$fstatistic
  p <- pf(f[1], f[2], f[3], lower.tail = FALSE)

  model <- "model: brain_swell ~ cerebral_hb_tot + hematocrit"
  r2 <- paste0("adjusted R^2: ", lm_fig4 %>% summary() %>% .$adj.r.squared %>% round(2))
  model_p_val <- paste0("model: ", round(p, 3))
  hb_tot_p_val <- paste0("cerebral_hb_tot: ", lm_fig4 %>% tidy() %>% .[2, "p.value"] %>% .$p.value %>% round(3))

# plot variables
require(ggpmisc)

df_fig4_outliers %>% 
  ggplot(aes(x = cerebral_hb_tot, y = brain_swell)) +
  geom_point(shape = 1, aes(size = hct, color = as.factor(brain_swell))) +
  #geom_abline(intercept = ) +

  geom_smooth(method = "lm", lty = 2, color = "#696969", alpha = 0.2) +
  # annotate("text", label = model, x = 25, y = 8, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = r2, x = 25, y = 7.5, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = "p-values", x = 25, y = 7, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = model_p_val, x = 35, y = 6.75, hjust = 0, vjust = 1, size = 4) +
  # annotate("text", label = hb_tot_p_val, x = 35, y = 6.5, hjust = 0, vjust = 1, size = 4) +

  lims(x = c(25, 225)) +
  theme_bw() +

  theme(# text = element_text(family = "Arial"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.text = element_text(size = 15),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

# 3d plot?  
# require(plotly)
#   x <- df_fig4_outliers$cerebral_hb_tot
#   y <- df_fig4_outliers$hematocrit
#   z <- df_fig4_outliers$brain_swell
#   
#   df_fig4_outliers %>% plot_ly(x = x, y = y, z = z, type = "scatter3d", mode = "markers", color = x)

```
**Figure 4.** model: brain_swell ~ cerebral_hb_tot + hematocrit; adjusted R^2: 0.25; model p-value: 0.007; cerebral_hb_tot p-value: 0.003

\newpage

### Fig 5. Explanatory model of how to interpret hemoglobin concentration

See manuscript draft (figure created in PowerPoint)

## Supplement
### SFig 1. DFA analysis illustration

#### A. Raw signal
```{r fig_s1a}

df_figs1a <- read.table("Data/signal_segments/Hb_tot/TM0001CM01.txt")

df_figs1a %>%

  ggplot(aes(x = Time)) +
  geom_line(aes(y = THC)) +

>>>>

---
date: "`r Sys.Date()`"
output: pdf_document
---

@@

# Load libraries
library(tidyverse)
library(readxl)
library(janitor)
library(extrafont)

@@

@@

# extract hb_tot and hb_oxy for the brain and muscle
df_fig1 <- df_hbox %>% 
  select(contains(c("hb_tot","hb_oxy")), -contains(c("alpha", "bsl"))) %>%
  ungroup() %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("tissue", "variable"), sep = "_", extra = "merge") %>% 
  mutate(variable = factor(variable, levels = c("hb_tot", "hb_oxy")))

@@

### Fig 3. Hb_oxy alpha and Hb_tot alpha in the brain and muscle
```{r fig3}

# extract hb_tot and hb_oxy for the brain and muscle
df_fig3 <- df_hbox %>% 
  select(contains("alpha2")) %>%
  ungroup() %>% 
  pivot_longer(2:ncol(.), names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("tissue", "variable"), sep = "_", extra = "merge") %>% 
  mutate(variable = str_remove(variable, "2")) %>% 
  mutate(variable = factor(variable, levels = c("hb_tot_alpha", "hb_oxy_alpha")))

# plot
df_fig3 %>%

  ggplot(aes(x = status, y = value)) +
  geom_violin(aes(color = status), fill = "white") +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = status), size = 0.2, width = 0.5) +
  facet_grid(variable ~ tissue, scales = "free_y") +

@@

@@
